### 2015-02-24 paper.tex
Created paper and state diagram

\title{\textbf{The Automatic Toilet Freshener \\ \large \emph{An technical odyssey}}}

\author{Bastiaan Weijers \\ 3256669 \and Colin Smits \\ 4075390}
\date{}
\documentclass[a4paper, 11pt]{article}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{pdflscape}
\usepackage{hyperref}

\begin{document}
\maketitle

\section{Introduction}

\section{Technical Overview}

%include state diagrams, circuit

\subsection{States}
\label{sub:state}
To start off with the technical details, we need to define states in which are system can be operating. These states must be defined according to the usage of a toilet. So, just after trying out all the circuits with the sensors, we started thinking about distinctive features of a toilet use.

First of all, you would have to open a door in order to get into the toilet. After that, you would maybe turn on the light in order to see where you are heading, making sure you do not miss the seat and fall down and/or need to clean the floor afterwards. These are the first two point that would suggest the usage of a toilet.

Secondly, you might need to open the lid in order to be able to sit down, or, in the case of a male user, you might need to put the seat up in order to have an increased target area. These are indications of the type of use, while a standing person would not attempt a "Number 2" usage (if you still have a normal functioning brain). Some people might close the lid again after usage, which would indicate they are done using the toilet, setting a time reference for our future system.

After using the toilet, you should be flushing and leaving the area. Therefore, you would open the door (if you did not close it \ldots) and turn off the light (if you actually turned it on \ldots). These are indications of the end of an usage, which would require the freshener to do it's job and clear the area of "bad smells".

\subsection{States of the Sensory}
In \S\hyperref[sub:state]{\ref{sub:state}} we defined the general use case of a toilet system. Now we need to turn these transitions into factors which we can measure using our sensors.

Starting off with the entrance, we can see that we could place a \textbf{magnetic sensor} in the door opening to measure the state the door is located in. Furthermore, we can measure the amount of light in a room using our \textbf{Light Dependent Resistor} (LDR). A sudden increase of light \emph{might} indicate a person entering the room. You are even more certain when the door is being opened and closed almost simultaneously, but still there is no guarantee. As we will see later on, you can never be sure what a person is doing inside the toilet (\ldots).

As we stated as well, a person might need to open the lid. If we were to place the \textbf{magnetic sensor} on the lid, we might be able to see whether the toilet is being used even more carefully. In order to be able to decide if a person is sitting or standing, we can use the \textbf{distance sensor}. In general, a standing person is further away from the wall behind the toilet then a sitting person.

Using these possible positions we head for the farewell of the toilet for now. We can measure whether a person is gone away from the toilet by using the \textbf{distance sensor} again. After finishing someone might close the lid (\textbf{magnetic sensor}) and walk through the door (\textbf{magnetic sensor} possibly) and turning off the light (\textbf{LDR}).

\subsection{Decisions, decisions}

We have set out all the possible uses for our sensors. In the end, we decided to do the following placements:
\begin{description}
\item[Distance Sensor] We place the \textit{distance sensor} on the wall behind the toilet to measure the distance between the wall and a person.
\item[Magnetic Sensor] ...
\item[LDR] We locate the \textit{LDR} close to the light source in the toilet to measure the on/off transition of the light.
\end{description}

\noindent After setting the locations, we are able to define our state diagram. However, we still need to make sure that the other components are used as well:
\begin{description}
\item[Temperature sensor] Located somewhere in the toilet just to measure the temperature. Is not used in deciding the states.
\item[Push Buttons] Two of the push buttons will be used for menu control, setting the delay of the spray as well as a reset for the spray counter. The other push button will be the manual override for a spray.
\item[LEDs and LCD Screen] All of these will be used as indicators for states and feedback towards the user.
\end{description}

Our final \hyperref[fig:stateDia]{state diagram} can be seen at the end of the paper.

\section{Development}
\subsection{First circuits}
In order to create a system that is able to do such a difficult task, we had to be able to create the circuits. As we were progressing through all the sensors, one of us hooked up his temperature sensor for testing purposes. It was then that we noticed a smell. The LCD display was not able to tell what it was, because it had no power. We pulled the USB cord from the computer immediately, as the smell was intensifying. After touching the temperature sensor, it became clear that this was the source: it was too hot too handle. This was the first piece which was harmed (but not destroyed), and in the full process of developing, some other pieces would fall victim to bad wiring.

\subsection{The final circuit}
After a week of intense coding and caffeine intake, along with some shocking progress on the circuit, we had created our final circuit without the refresher hooked up yet. We found that our code was fairly able to define our states, so we decided to give it a try. It was then we victimized the second part of our kit: after an hour of trying to get the refresher to work on pin 13, we found that the circuit didn't work, as the led was on, but there was no current going through the refresher. So we decided to replace the parts one by one, only to find out that the MOSFET was not working any more. It was pronounced dead 20th of February, around 6 pm. We then had our circuit fully working, as can be seen in \hyperref[fig:circuit]{this figure}.

To make sure we were able to incorporate all the parts, we have decided to put the push buttons on the analogue inputs as it was fairly easy to code the buttons from there as well. Along with that, we had to put two of the LEDs on the analogue inputs due to lack op space. The decision was made as well to hook the refresher onto pin 13, so that the LED on the Arduino itself would be activated in line with the refresher, so it's easy to see when the refresher is active.

\section{Usage}

\section{Reflection}

%state diagram in landscape mode
\begin{figure}
\includegraphics[scale=0.5]{schakeling_bb.png}
\label{fig:circuit}
\caption{The breadboard circuit of our system}
\end{figure}  
\newpage
\begin{landscape}
\begin{figure}[h!]
\includegraphics[scale=0.7]{State_Diagram.pdf}
\label{fig:stateDia}
\caption{State Diagram of our system}
\end{figure}
\end{landscape}

\end{document}

### 2015-03-17 pract2/test.py
inserted control sequence for comm with win

@@ -1,7 +1,7 @@
 import numpy as np 
 import cv2
 
-img = cv2.imread('oefen.jpg', 0)
+img = cv2.imread('oefen.jpg', -1)
 cv2.imshow('image', img)
 cv2.waitKey(0)
 cv2.destroyAllWindows()

### 2015-03-17 pract2/testBackground.py

@@ -1,0 +1,16 @@
+import numpy as np
+import cv2
+
+imgBack = cv2.imread("background.jpg", 0)
+imgFore = cv2.imread("inserted.jpg", 0)
+
+fgbg = cv2.BackgroundSubtractorMOG()
+
+fgmask = fgbg.apply(imgBack)
+bgmask = fgbg.apply(imgFore)
+
+cv2.imshow('frame', fgmask)
+cv2.waitKey(0)
+cv2.imshow('frame', bgmask)
+cv2.waitKey(0)
+cv2.destroyAllWindows()

### 2015-03-17 pract2/testFace.py

@@ -1,0 +1,22 @@
+import numpy as np
+import cv2
+
+face_cascade = cv2.CascadeClassifier('haarcascade_frontalface_default.xml')
+eye_cascade = cv2.CascadeClassifier('haarcascade_eye.xml')
+
+img = cv2.imread('face.jpg')
+gray = cv2.cvtColor(img, cv2.COLOR_BGR2GRAY)
+
+
+faces = face_cascade.detectMultiScale(gray, 1.3, 5)
+for (x,y,w,h) in faces:
+    cv2.rectangle(img,(x,y),(x+w,y+h),(255,0,0),2)
+    roi_gray = gray[y:y+h, x:x+w]
+    roi_color = img[y:y+h, x:x+w]
+    eyes = eye_cascade.detectMultiScale(roi_gray)
+    for (ex,ey,ew,eh) in eyes:
+        cv2.rectangle(roi_color,(ex,ey),(ex+ew,ey+eh),(0,255,0),2)
+
+cv2.imshow('img',img)
+cv2.waitKey(0)
+cv2.destroyAllWindows()

### 2015-03-17 pract2/testOpticalFlow.py

@@ -1,0 +1,35 @@
+import cv2
+import numpy as np
+
+cap = cv2.VideoCapture(0)
+while (True):
+    ret, frame1 = cap.read()
+    if (not frame1 == None):
+        break
+
+prvs = cv2.cvtColor(frame1, cv2.COLOR_BGR2GRAY)
+hsv = np.zeros_like(frame1)
+hsv[...,1] = 255
+
+while(1):
+    ret, frame2 = cap.read()
+    next = cv2.cvtColor(frame2,cv2.COLOR_BGR2GRAY)
+
+    flow = cv2.calcOpticalFlowFarneback(prvs,next, 0.5, 3, 15, 3, 5, 1.2, 0)
+
+    mag, ang = cv2.cartToPolar(flow[...,0], flow[...,1])
+    hsv[...,0] = ang*180/np.pi/2
+    hsv[...,2] = cv2.normalize(mag,None,0,255,cv2.NORM_MINMAX)
+    rgb = cv2.cvtColor(hsv,cv2.COLOR_HSV2BGR)
+
+    cv2.imshow('frame2',rgb)
+    k = cv2.waitKey(30) & 0xff
+    if k == 27:
+        break
+    elif k == ord('s'):
+        cv2.imwrite('opticalfb.png',frame2)
+        cv2.imwrite('opticalhsv.png',rgb)
+    prvs = next
+
+cap.release()
+cv2.destroyAllWindows()

### 2015-03-17 pract2/testVideoBackground.py

@@ -1,0 +1,19 @@
+import numpy as np
+import cv2
+
+cap = cv2.VideoCapture(0)
+
+fgbg = cv2.BackgroundSubtractorMOG()
+
+while(cap.isOpened):
+    ret, frame = cap.read()
+
+    if ret == True:
+    	fgmask = fgbg.apply(frame)
+        cv2.imshow('frame',fgmask)
+    k = cv2.waitKey(30) & 0xff
+    if k == 27:
+       break
+
+cap.release()
+cv2.destroyAllWindows()

### 2015-03-17 pract2/tryOutCom.py

@@ -1,0 +1,21 @@
+import win32com.client as comclt
+import win32api
+
+# Initialiseer de Windows shell in Python zodat je naar de shell kunt scripten
+wsh= comclt.Dispatch("WScript.Shell")
+
+
+
+# Voor elke keer dat je een opdracht wilt versturen:
+wsh.AppActivate("Photo Gallery") # select another application
+wsh.SendKeys("^.")
+
+
+"""
+SendKeys: next Photo   = wsh.SendKeys("{Right}")
+          prev Photo   = wsh.SendKeys("{Left}")
+          Rotate left  = wsh.SendKeys("^.")
+          Rotate Right = wsh.SendKeys("^,")
+
+"""
+

### 2015-03-17 pract2/center.py
some new files
The following have been added:
- functions, containing video manipulation code (detecting hands,
drawing)
- jpeg files for testing
- testBackground: tracking blobs in video
- center: showing arrow from detected objects in 2 pictures

@@ -1,0 +1,37 @@
+import numpy as np
+import cv2
+import functions
+
+img1 = cv2.imread("testOtherLighting.jpg")
+img2 = cv2.imread("test4OtherLighting.jpg")
+
+fgbg = cv2.BackgroundSubtractorMOG()
+
+
+#frame = cv2.imread("test2OtherLighting.jpg")
+#skin filter boundaries (tweak if necessary)
+lower = np.array([0, 45, 80], dtype = "uint8")
+upper = np.array([11, 255, 255], dtype = "uint8")
+
+list1 = functions.skinFilter(img1, lower, upper)
+img1 = list1[0]
+center1 = list1[1]
+cv2.imshow("image1", img1)  
+
+list2 = functions.skinFilter(img2, lower, upper)
+img2 = list2[0]
+center2 = list2[1]
+
+
+cv2.arrowedLine(img2,center1, center2,[0,255,255],2)
+cv2.imshow("image2", img2)   
+
+
+cv2.waitKey(0)
+
+cv2.destroyAllWindows()
+
+
+
+
+

### 2015-03-17 pract2/functions.py

@@ -1,0 +1,74 @@
+import cv2
+import numpy as np
+
+def skinFilter(frame, lower, upper):
+	converted = cv2.cvtColor(frame, cv2.COLOR_BGR2HSV)
+	skinMask = cv2.inRange(converted, lower, upper)
+	kernel = cv2.getStructuringElement(cv2.MORPH_ELLIPSE, (11, 11))
+	skinMask = cv2.erode(skinMask, kernel, iterations = 3)
+	skinMask = cv2.dilate(skinMask, kernel, iterations = 3)
+	skinMask = cv2.GaussianBlur(skinMask, (3,3), 0)
+	skin = cv2.bitwise_and(frame, frame, mask = skinMask)
+	skin = cv2.cvtColor(skin, cv2.COLOR_RGB2GRAY)
+	ret, tresh1 = cv2.threshold(skin, 0,255,cv2.THRESH_BINARY)
+	center = findDefects(frame, tresh1)
+	
+	return [frame, center]
+
+
+def findDefects(frame, gray):
+	contours,hierarchy = cv2.findContours(gray, 1, 2)
+	maxArea = 0
+	ci = 0
+	
+	n = 0
+	if (contours):
+		for i in range(len(contours)):
+			cnt = contours[n]
+			#if cv2.isContourConvex(cnt):
+			print "True"
+			print "\n"
+			area = cv2.contourArea(cnt)
+			print area
+			print "\n"
+			if area > maxArea:
+				maxArea = area
+				ci = n
+
+			n += 1
+		
+		#print str(contours)
+		cnt = contours[ci]
+		
+		M = cv2.moments(cnt)
+		cx = int(M['m10']/M['m00'])
+		cy = int(M['m01']/M['m00'])
+		#print len(contours)
+		#print ci
+		#print str(cnt)
+		hull = cv2.convexHull(cnt,returnPoints = False)
+		#print str(hull)
+		defects = cv2.convexityDefects(cnt,hull)
+		#print str(defects)
+
+		for i in range(defects.shape[0]):
+			s,e,f,d = defects[i,0]
+			start = tuple(cnt[s][0])
+			end = tuple(cnt[e][0])
+			far = tuple(cnt[f][0])
+			#print start
+			#print end
+			#print far
+			cv2.line(frame,start,end,[0,255,0],2)
+			cv2.circle(frame,far,5,[0,0,255],-1)
+
+	return (cx, cy)
+			
+
+
+
+
+
+
+
+

### 2015-03-17 pract2/testBackground.py

@@ -1,16 +1,50 @@
 import numpy as np
 import cv2
+import functions
 
-imgBack = cv2.imread("background.jpg", 0)
-imgFore = cv2.imread("inserted.jpg", 0)
-
+cap = cv2.VideoCapture(0)
 fgbg = cv2.BackgroundSubtractorMOG()
 
-fgmask = fgbg.apply(imgBack)
-bgmask = fgbg.apply(imgFore)
 
-cv2.imshow('frame', fgmask)
+#frame = cv2.imread("test2OtherLighting.jpg")
+#skin filter boundaries (tweak if necessary)
+lower = np.array([0, 45, 80], dtype = "uint8")
+upper = np.array([11, 255, 255], dtype = "uint8")
+
+
+
+#"""
+#starting cam: no frame yet
+while (True):
+    ret, frame1 = cap.read()
+    if (not frame1 == None):
+        break
+
+#cam started, process frames
+while(cap.isOpened):
+    ret, frame = cap.read()
+
+    if ret == True:
+#"""
+
+		img = frame.copy()
+		#back = fgbg.apply(img)
+		img = functions.skinFilter(img, lower, upper)
+		cv2.imshow("images", img)
+
+
+#"""      
+    k = cv2.waitKey(30) & 0xff
+    if k == 27:
+       break
+
+cap.release()
+"""
 cv2.waitKey(0)
-cv2.imshow('frame', bgmask)
-cv2.waitKey(0)
+"""
 cv2.destroyAllWindows()
+
+
+
+
+

### 2015-03-17 pract2/testFace.py

@@ -4,19 +4,42 @@ import cv2
 face_cascade = cv2.CascadeClassifier('haarcascade_frontalface_default.xml')
 eye_cascade = cv2.CascadeClassifier('haarcascade_eye.xml')
 
-img = cv2.imread('face.jpg')
-gray = cv2.cvtColor(img, cv2.COLOR_BGR2GRAY)
+cap = cv2.VideoCapture(0)
 
 
-faces = face_cascade.detectMultiScale(gray, 1.3, 5)
-for (x,y,w,h) in faces:
-    cv2.rectangle(img,(x,y),(x+w,y+h),(255,0,0),2)
-    roi_gray = gray[y:y+h, x:x+w]
-    roi_color = img[y:y+h, x:x+w]
-    eyes = eye_cascade.detectMultiScale(roi_gray)
-    for (ex,ey,ew,eh) in eyes:
-        cv2.rectangle(roi_color,(ex,ey),(ex+ew,ey+eh),(0,255,0),2)
 
-cv2.imshow('img',img)
+#starting cam: no frame yet
+while (True):
+    ret, frame1 = cap.read()
+    if (not frame1 == None):
+        break
+
+#cam started, process frames
+while(cap.isOpened):
+    ret, frame = cap.read()
+
+    if ret == True:
+		img = frame.copy()
+		gray = cv2.cvtColor(img, cv2.COLOR_BGR2GRAY)
+		faces = face_cascade.detectMultiScale(gray, 1.3, 5)
+		print str(faces)
+		for (x,y,w,h) in faces:
+		    cv2.rectangle(img,(x,y),(x+w,y+h),(255,0,0),2)
+		    roi_gray = gray[y:y+h, x:x+w]
+		    roi_color = img[y:y+h, x:x+w]
+		    eyes = eye_cascade.detectMultiScale(roi_gray)
+		    for (ex,ey,ew,eh) in eyes:
+		        cv2.rectangle(roi_color,(ex,ey),(ex+ew,ey+eh),(0,255,0),2)
+
+		cv2.imshow('img',gray)
+     
+    k = cv2.waitKey(30) & 0xff
+    if k == 27:
+       break
+
+cap.release()
+"""
 cv2.waitKey(0)
+"""
 cv2.destroyAllWindows()
+

### 2015-03-18 pract2/functions.py
First working prototype
Swipen werkt voor iemand met korte mouwen, halve meter afstand

@@ -4,13 +4,15 @@ import numpy as np
 def skinFilter(frame, lower, upper):
 	converted = cv2.cvtColor(frame, cv2.COLOR_BGR2HSV)
 	skinMask = cv2.inRange(converted, lower, upper)
-	kernel = cv2.getStructuringElement(cv2.MORPH_ELLIPSE, (11, 11))
-	skinMask = cv2.erode(skinMask, kernel, iterations = 3)
-	skinMask = cv2.dilate(skinMask, kernel, iterations = 3)
+	kernel = cv2.getStructuringElement(cv2.MORPH_ELLIPSE, (10, 10))
+	skinMask = cv2.erode(skinMask, kernel, iterations = 2)
+	skinMask = cv2.dilate(skinMask, kernel, iterations = 2)
 	skinMask = cv2.GaussianBlur(skinMask, (3,3), 0)
 	skin = cv2.bitwise_and(frame, frame, mask = skinMask)
 	skin = cv2.cvtColor(skin, cv2.COLOR_RGB2GRAY)
 	ret, tresh1 = cv2.threshold(skin, 0,255,cv2.THRESH_BINARY)
+	cv2.imshow("binary", tresh1)
+
 	center = findDefects(frame, tresh1)
 	
 	return [frame, center]
@@ -22,7 +24,9 @@ def findDefects(frame, gray):
 	ci = 0
 	
 	n = 0
-	if (contours):
+	cx = None
+	cy = None
+	if len(contours) > 1:
 		for i in range(len(contours)):
 			cnt = contours[n]
 			#if cv2.isContourConvex(cnt):
@@ -40,35 +44,51 @@ def findDefects(frame, gray):
 		#print str(contours)
 		cnt = contours[ci]
 		
-		M = cv2.moments(cnt)
-		cx = int(M['m10']/M['m00'])
-		cy = int(M['m01']/M['m00'])
-		#print len(contours)
-		#print ci
-		#print str(cnt)
-		hull = cv2.convexHull(cnt,returnPoints = False)
-		#print str(hull)
-		defects = cv2.convexityDefects(cnt,hull)
-		#print str(defects)
+		if maxArea > 10000:
+			M = cv2.moments(cnt)
+			cx = int(M['m10']/M['m00'])
+			cy = int(M['m01']/M['m00'])
+			#print len(contours)
+			#print ci
+			#print str(cnt)
+			hull = cv2.convexHull(cnt,returnPoints = False)
+			#print str(hull)
[…]

### 2015-03-18 pract2/testBackground.py

@@ -2,6 +2,13 @@ import numpy as np
 import cv2
 import functions
 
+import win32com.client as comclt
+import win32api
+
+# Initialiseer de Windows shell in Python zodat je naar de shell kunt scripten
+wsh= comclt.Dispatch("WScript.Shell")
+
+
 cap = cv2.VideoCapture(0)
 fgbg = cv2.BackgroundSubtractorMOG()
 
@@ -9,8 +16,8 @@ fgbg = cv2.BackgroundSubtractorMOG()
 #frame = cv2.imread("test2OtherLighting.jpg")
 #skin filter boundaries (tweak if necessary)
 lower = np.array([0, 45, 80], dtype = "uint8")
-upper = np.array([11, 255, 255], dtype = "uint8")
-
+upper = np.array([15, 255, 255], dtype = "uint8")
+handFound = False
 
 
 #"""
@@ -20,17 +27,66 @@ while (True):
     if (not frame1 == None):
         break
 
+lastFrame = 0
+lastFound = 0
+findDelay = 20
+maxDelay = 20
+
+trigger = 400
+
 #cam started, process frames
 while(cap.isOpened):
     ret, frame = cap.read()
 
     if ret == True:
-#"""
 
-		img = frame.copy()
-		#back = fgbg.apply(img)
-		img = functions.skinFilter(img, lower, upper)
-		cv2.imshow("images", img)
+        if lastFound < findDelay:
+            lastFound += 1
+
+        img = frame.copy()
+       # back = fgbg.apply(img)
+       # img = cv2.bitwise_and(img, img, mask=back)
+       # kernel = cv2.getStructuringElement(cv2.MORPH_ELLIPSE, (3, 3))
+       # img = cv2.erode(img, kernel, iterations = 2)
+       # img = cv2.dilate(img, kernel, iterations = 2)
+        listImg = functions.skinFilter(img, lower, upper)
+        img = listImg[0]
+        center = listImg[1]
+        cv2.imshow("images", img)
+
+        if (center[0] != None):
+            if (handFound == True):
+                prev_Offset = functions.processHand(center, prev_center, prev_Offset)
+                
+                length = functions.getLengthX(prev_Offset)
+                if abs(length) > trigger:
+                    print length
+                    if length < 0:
+                        wsh.AppActivate("Photo Gallery") # select another application
+                        wsh.sendKeys("{Left}")
+                    else:
+                        wsh.AppActivate("Photo Gallery")
+                        wsh.sendKeys("{Right}")
+
[…]

### 2015-03-18 pract2/functions.py
werkend rotatie en swipe
Vereisten:
- lange mouwen
- gezicht uit scherm
- handen uitgestrekt (geen vuist)
nog bekijken:
- achtergrond niet getest (werkt op huidskleur)
-

@@ -13,9 +13,9 @@ def skinFilter(frame, lower, upper):
 	ret, tresh1 = cv2.threshold(skin, 0,255,cv2.THRESH_BINARY)
 	cv2.imshow("binary", tresh1)
 
-	center = findDefects(frame, tresh1)
+	centers = findDefects(frame, tresh1)
 	
-	return [frame, center]
+	return [frame, centers]
 
 
 def findDefects(frame, gray):
@@ -26,6 +26,8 @@ def findDefects(frame, gray):
 	n = 0
 	cx = None
 	cy = None
+	ci = []
+	centers = []
 	if len(contours) > 1:
 		for i in range(len(contours)):
 			cnt = contours[n]
@@ -35,19 +37,18 @@ def findDefects(frame, gray):
 			area = cv2.contourArea(cnt)
 			print area
 			print "\n"
-			if area > maxArea:
-				maxArea = area
-				ci = n
+			if area > 10000:
+				ci.append(i)
 
 			n += 1
 		
 		#print str(contours)
-		cnt = contours[ci]
-		
-		if maxArea > 10000:
+		for i in ci:
+			cnt = contours[i]
 			M = cv2.moments(cnt)
 			cx = int(M['m10']/M['m00'])
 			cy = int(M['m01']/M['m00'])
+			centers.append((cx, cy))
 			#print len(contours)
 			#print ci
 			#print str(cnt)
@@ -67,11 +68,11 @@ def findDefects(frame, gray):
 				cv2.line(frame,start,end,[0,255,0],2)
 				cv2.circle(frame,far,5,[0,0,255],-1)
 
-	return (cx, cy)
+	return centers
 			
 
 def processHand(center, prev_center, prev_Offset):
-	
+
 
 	difference_X = center[0] - prev_center[0]
 	difference_y = center[1] - prev_center[1]
@@ -92,3 +93,24 @@ def getLengthX(vector):
 
 def getLengthY(vector):
 	return vector[1]
+
+
+def getPreviousCenter(center, previous_centers):
+	i = 0
+	maxDistance = 10000000
+	
+	for prev_center in previous_centers:
+
+		difference_X = center[0] - prev_center[0]
+		difference_Y = center[1] - prev_center[1]
+
+		totalDifference = abs(difference_X)
+
+		if totalDifference < maxDistance:
+			maxDistance = totalDifference
+			current = i
+
+		i += 1
+
+	return current	
+

### 2015-03-18 pract2/testBackground.py

@@ -28,11 +28,17 @@ while (True):
         break
 
 lastFrame = 0
+lastFrame2 = 0
 lastFound = 0
 findDelay = 20
 maxDelay = 20
+oneHand = False
+twoHands = False
+trigger = 300
+_2HandsFound = False
+lastFound2 = 0
 
-trigger = 400
+trigger2Hands = 200
 
 #cam started, process frames
 while(cap.isOpened):
@@ -43,6 +49,14 @@ while(cap.isOpened):
         if lastFound < findDelay:
             lastFound += 1
 
+        if lastFound2 < findDelay:
+            lastFound2 += 1
+
+        if lastFrame2 < findDelay:
+            lastFrame2 += 1
+        else:
+            twoHands = False
+
         img = frame.copy()
        # back = fgbg.apply(img)
        # img = cv2.bitwise_and(img, img, mask=back)
@@ -51,13 +65,16 @@ while(cap.isOpened):
        # img = cv2.dilate(img, kernel, iterations = 2)
         listImg = functions.skinFilter(img, lower, upper)
         img = listImg[0]
-        center = listImg[1]
+        centers = listImg[1]
         cv2.imshow("images", img)
 
-        if (center[0] != None):
+        
+        if len(centers) == 1 and not twoHands:
+            print len(centers)
+            print centers
             if (handFound == True):
-                prev_Offset = functions.processHand(center, prev_center, prev_Offset)
-                
+                prev_Offset = functions.processHand(centers[0], prev_centers[0], prev_Offset)
+                prev_centers = centers
                 length = functions.getLengthX(prev_Offset)
                 if abs(length) > trigger:
                     print length
@@ -75,15 +92,97 @@ while(cap.isOpened):
             elif lastFound >= findDelay:
                 handFound = True
                 prev_Offset = (0,0)
-                prev_center = center
+                prev_centers = centers
 
             lastFrame = 0
+            oneHand = True
+
+
+
+        elif len(centers) == 2:
+            twoHands = True
+            related = []
+            lengths = dict()
+            if (_2HandsFound == True):
+                for center in centers:
+                    print len(prev_centers)
+                    related.append(functions.getPreviousCenter(center, prev_centers))
+                    print related
+
+                if related[0] == related[1]:
+                    print "Kan niet"
+
+                if related[0] == 1:
+                    centers.reverse()
+
+                print centers
+                for i in range(len(centers)):
+                    print related[i]
+                    print len(centers)
[…]

### 2015-03-19 pract2/faceDetect.py
try ip config

@@ -1,0 +1,40 @@
+import numpy as np
+import cv2
+
+face_cascade = cv2.CascadeClassifier('casc.xml')
+
+
+cap = cv2.VideoCapture()
+if not cap.open("http://192.168.173.176:8080"):
+	print "Didn't work"
+
+#starting cam: no frame yet
+while (True):
+    ret, frame1 = cap.read()
+    if (not frame1 == None):
+        break
+
+while(cap.isOpened):
+	ret, frame = cap.read()
+
+	if ret == True:
+		img = frame.copy()
+		# gray = cv2.cvtColor(img, cv2.COLOR_BGR2GRAY)
+		# faces = face_cascade.detectMultiScale(gray, 1.1 , 5, minSize=(30, 30), flags=cv2.cv.CV_HAAR_SCALE_IMAGE)
+		# print len(faces)
+		# for (x,y,w,h) in faces:
+		#     cv2.rectangle(img,(x,y),(x+w,y+h),(255,0,0),2)
+		#     # roi_gray = gray[y:y+h, x:x+w]
+		#     # roi_color = img[y:y+h, x:x+w]
+		#     # eyes = eye_cascade.detectMultiScale(roi_gray)
+		#     # for (ex,ey,ew,eh) in eyes:
+		#     #     cv2.rectangle(roi_color,(ex,ey),(ex+ew,ey+eh),(0,255,0),2)
+
+		cv2.imshow('img',img)
+
+	k = cv2.waitKey(30) & 0xff
+	if k == 27:
+		break
+
+cap.release()
+cv2.destroyAllWindows()

### 2015-03-19 pract2/paper.tex
newest updated working version

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Bastiaan Weijers \\ 3256669 \and Colin Smits \\4075390}
\title{Body Photo Browser}
\renewcommand*{\familydefault}{\sfdefault}
\begin{document}
\maketitle

\section{Introduction}
Currently, virtual reality is starting to grow in the digital world, which makes it possible for a user to move in a virtual world using motion, which is then detected by sensors. This paper will discuss an implementation of processing hand gestures in order to operate a photo browser system (Windows Photo Gallery\footnote{http://windows.microsoft.com/nl-nl/windows-live/photo-gallery}). First, we will discuss the techniques used in order to create a working system, and after that the requirements needed for the system to work will be explained.

\section{Processing Steps}
For building our system, we used Python as programming language, due to it's ease of use with matrix calculations. The testing was done using a 1.5MP integrated web cam. Our system is based on detecting the hands using color filters, and getting the center of the hands which can then be tracked. The following gestures can be recognised:

\begin{description}
\item[Swipe Left:] Using one hand (in view), going from right to left in the view. This will give the previous photo.
\item[Swipe Right:] Using one hand (in view), going from left to right in the view. This will give the next photo.
\item[Rotate left:] Using two hands (in view), going up with the right hand, and simultaneously going down with the left hand. This mimics the idea of rotating a picture counter clockwise.
\item[Rotate Right:] Using two hands (in view), going up with the left hand, and simultaneously going down with the right hand. This quite mimics the idea of rotating a picture clockwise.
\end{description}

Firstly, we process the frame by converting it from BGR (channels switched)into a YCrCb color coded file. Then, we apply the cv.inRange() function in order to filter the pixels that are in the color range. Then, we use 2 iterations of erosion and dilation, and a Gaussian blur, after which we compare the original image to the \emph{mask}. This generates a better qualitative comparison, as well as bigger area, so that contours can be detected more easily (explained further on). The pixels which are contained in both files will be included in the further processed file.\\
\indent Secondly, we try to detect hands in the picture. For that we use contours by using the OpenCV operation findContours(), after which we filter contours with an area too small, as these may bbe noise inducted contours. We then find the center of each large enough contour by applying cv.moments(). Using these centers, we can track the hands and recognise gestures by comparing the center's location per frame.
\section{Assumptions \& Requirements}
In order for our system to work we have to set the following requirements:

\begin{itemize}
\item Due to the fact we detect skin color, the only parts included in the view must be hands. If a face is in view, it might be detected as a contour, and thus be classified as hand.
\item As we want to detect hands, the user must wear long sleeves. If there is a "naked" arm in view, the moments function will find a COG somewhere in the middle of the arm. This COG will have less movement, so nothing is detected.
\item The background must be of a color which is not similar to the user's skin color, i.e. a wooden chest will not suffice as a valid background. A white wall, however, will.
\item Due to color boundary issues, the user must be a person having a light skin color.

\end{itemize}

\section{Result}

\end{document}

### 2015-03-19 pract2/bodyMotion.py

@@ -15,8 +15,8 @@ fgbg = cv2.BackgroundSubtractorMOG()
 
 #frame = cv2.imread("test2OtherLighting.jpg")
 #skin filter boundaries (tweak if necessary)
-lower = np.array([0, 45, 80], dtype = "uint8")
-upper = np.array([15, 255, 255], dtype = "uint8")
+lower = np.array([0, 144, 77], dtype = "uint8")
+upper = np.array([255, 173, 127], dtype = "uint8")
 handFound = False
 
 
@@ -30,11 +30,11 @@ while (True):
 lastFrame = 0
 lastFrame2 = 0
 lastFound = 0
-findDelay = 20
+findDelay = 30
 maxDelay = 20
 oneHand = False
 twoHands = False
-trigger = 300
+trigger = 250
 _2HandsFound = False
 lastFound2 = 0
 
@@ -56,28 +56,35 @@ while(cap.isOpened):
             lastFrame2 += 1
         else:
             twoHands = False
+            _2HandsFound = False
 
         img = frame.copy()
-       # back = fgbg.apply(img)
-       # img = cv2.bitwise_and(img, img, mask=back)
-       # kernel = cv2.getStructuringElement(cv2.MORPH_ELLIPSE, (3, 3))
-       # img = cv2.erode(img, kernel, iterations = 2)
-       # img = cv2.dilate(img, kernel, iterations = 2)
+        """
+        back = fgbg.apply(img)
+        img = cv2.bitwise_and(img, img, mask=back)
+        kernel = cv2.getStructuringElement(cv2.MORPH_ELLIPSE, (3, 3))
+        img = cv2.erode(img, kernel, iterations = 2)
+        img = cv2.dilate(img, kernel, iterations = 2)
+        """
         listImg = functions.skinFilter(img, lower, upper)
         img = listImg[0]
         centers = listImg[1]
-        cv2.imshow("images", img)
+        
+        
+       
+        cv2.imshow("img", img)
+        
 
         
+    
         if len(centers) == 1 and not twoHands:
-            print len(centers)
-            print centers
+            
             if (handFound == True):
                 prev_Offset = functions.processHand(centers[0], prev_centers[0], prev_Offset)
                 prev_centers = centers
                 length = functions.getLengthX(prev_Offset)
                 if abs(length) > trigger:
-                    print length
+                    #print length
                     if length < 0:
                         wsh.AppActivate("Photo Gallery") # select another application
[…]

### 2015-03-19 pract2/functions.py

@@ -2,7 +2,7 @@ import cv2
 import numpy as np
 
 def skinFilter(frame, lower, upper):
-	converted = cv2.cvtColor(frame, cv2.COLOR_BGR2HSV)
+	converted = cv2.cvtColor(frame, cv2.COLOR_BGR2YCR_CB)
 	skinMask = cv2.inRange(converted, lower, upper)
 	kernel = cv2.getStructuringElement(cv2.MORPH_ELLIPSE, (10, 10))
 	skinMask = cv2.erode(skinMask, kernel, iterations = 2)
@@ -31,40 +31,27 @@ def findDefects(frame, gray):
 	if len(contours) > 1:
 		for i in range(len(contours)):
 			cnt = contours[n]
-			#if cv2.isContourConvex(cnt):
-			print "True"
-			print "\n"
 			area = cv2.contourArea(cnt)
-			print area
-			print "\n"
+					
 			if area > 10000:
 				ci.append(i)
 
 			n += 1
 		
-		#print str(contours)
 		for i in ci:
 			cnt = contours[i]
 			M = cv2.moments(cnt)
 			cx = int(M['m10']/M['m00'])
 			cy = int(M['m01']/M['m00'])
 			centers.append((cx, cy))
-			#print len(contours)
-			#print ci
-			#print str(cnt)
 			hull = cv2.convexHull(cnt,returnPoints = False)
-			#print str(hull)
 			defects = cv2.convexityDefects(cnt,hull)
-			#print str(defects)
 
 			for i in range(defects.shape[0]):
 				s,e,f,d = defects[i,0]
 				start = tuple(cnt[s][0])
 				end = tuple(cnt[e][0])
 				far = tuple(cnt[f][0])
-				#print start
-				#print end
-				#print far
 				cv2.line(frame,start,end,[0,255,0],2)
 				cv2.circle(frame,far,5,[0,0,255],-1)
 
@@ -98,7 +85,8 @@ def getLengthY(vector):
 def getPreviousCenter(center, previous_centers):
 	i = 0
 	maxDistance = 10000000
-	
+	print "Previous: %s" % (previous_centers)
+	print "Current: %s" % (str(center))
 	for prev_center in previous_centers:
 
 		difference_X = center[0] - prev_center[0]
@@ -111,6 +99,8 @@ def getPreviousCenter(center, previous_centers):
 			current = i
 
 		i += 1
+
+	print "Related to center %s: %s = %s" % (str(center), current, previous_centers[current])
 
 	return current	
 

### 2015-03-19 pract2/bodyMotion.py
commented

@@ -5,22 +5,20 @@ import functions
 import win32com.client as comclt
 import win32api
 
-# Initialiseer de Windows shell in Python zodat je naar de shell kunt scripten
+# Initialise Windows shell for communication
 wsh= comclt.Dispatch("WScript.Shell")
 
 
 cap = cv2.VideoCapture(0)
-fgbg = cv2.BackgroundSubtractorMOG()
+#fgbg = cv2.BackgroundSubtractorMOG()
 
-
-#frame = cv2.imread("test2OtherLighting.jpg")
 #skin filter boundaries (tweak if necessary)
 lower = np.array([0, 144, 77], dtype = "uint8")
 upper = np.array([255, 173, 127], dtype = "uint8")
 handFound = False
 
 
-#"""
+
 #starting cam: no frame yet
 while (True):
     ret, frame1 = cap.read()
@@ -46,6 +44,7 @@ while(cap.isOpened):
 
     if ret == True:
 
+        # Delays in order to make sure a gesture will not be processed twice at once
         if lastFound < findDelay:
             lastFound += 1
 
@@ -58,7 +57,9 @@ while(cap.isOpened):
             twoHands = False
             _2HandsFound = False
 
+
         img = frame.copy()
+
         """
         back = fgbg.apply(img)
         img = cv2.bitwise_and(img, img, mask=back)
@@ -66,7 +67,12 @@ while(cap.isOpened):
         img = cv2.erode(img, kernel, iterations = 2)
         img = cv2.dilate(img, kernel, iterations = 2)
         """
+
+        # process the image to find skin locations
+        # Returns list: [ img, centers ] where centers = [(int, int)]
         listImg = functions.skinFilter(img, lower, upper)
+
+
         img = listImg[0]
         centers = listImg[1]
         
@@ -75,8 +81,7 @@ while(cap.isOpened):
         cv2.imshow("img", img)
         
 
-        
-    
+        # one hand found 
         if len(centers) == 1 and not twoHands:
             
             if (handFound == True):
@@ -84,28 +89,34 @@ while(cap.isOpened):
                 prev_centers = centers
                 length = functions.getLengthX(prev_Offset)
                 if abs(length) > trigger:
-                    #print length
-                    if length < 0:
-                        wsh.AppActivate("Photo Gallery") # select another application
+                    # put Photo Gallery up as active
+                    # send the right key 
+                    if length < 0: 
+                        # previous
+                        wsh.AppActivate("Photo Gallery")
                         wsh.sendKeys("{Left}")
[…]

### 2015-03-19 pract2/functions.py

@@ -1,24 +1,41 @@
 import cv2
 import numpy as np
 
+
+# Function to detect skin patches and identify their location
 def skinFilter(frame, lower, upper):
+
+	# convert frame to YCrCb from BGR (reversed RGB)
 	converted = cv2.cvtColor(frame, cv2.COLOR_BGR2YCR_CB)
+
+	#filter the pixels within the skin color range
 	skinMask = cv2.inRange(converted, lower, upper)
+
+	# process the frame (erosion, dilation) for better results
 	kernel = cv2.getStructuringElement(cv2.MORPH_ELLIPSE, (10, 10))
 	skinMask = cv2.erode(skinMask, kernel, iterations = 2)
 	skinMask = cv2.dilate(skinMask, kernel, iterations = 2)
 	skinMask = cv2.GaussianBlur(skinMask, (3,3), 0)
+
+	# Use AND operator to identify pixels that can be used
 	skin = cv2.bitwise_and(frame, frame, mask = skinMask)
+
+	#Threshold and create binary image for processing
 	skin = cv2.cvtColor(skin, cv2.COLOR_RGB2GRAY)
 	ret, tresh1 = cv2.threshold(skin, 0,255,cv2.THRESH_BINARY)
 	cv2.imshow("binary", tresh1)
 
+	#calculate the centers of the skin patches by identifying contours
 	centers = findDefects(frame, tresh1)
 	
 	return [frame, centers]
 
 
+
+# Function to find the defects and contours in the frame
 def findDefects(frame, gray):
+
+	# find the contours
 	contours,hierarchy = cv2.findContours(gray, 1, 2)
 	maxArea = 0
 	ci = 0
@@ -29,6 +46,8 @@ def findDefects(frame, gray):
 	ci = []
 	centers = []
 	if len(contours) > 1:
+
+		# for each contour: calculate the area and filter the patches that are too small
 		for i in range(len(contours)):
 			cnt = contours[n]
 			area = cv2.contourArea(cnt)
@@ -38,6 +57,13 @@ def findDefects(frame, gray):
 
 			n += 1
 		
+		# for each contour left:
+		# - calculate the COG by using the moments of the contour (http)
+		# - Get the defects of the fingers
+		# - By defining the defect location: draw lines between the fingertips
+		# 	creating a contour of the hand
+		# - Draw the contour of the hand on the frame
+
 		for i in ci:
 			cnt = contours[i]
 			M = cv2.moments(cnt)
[…]

### 2015-03-25 pract2/paper.tex
Start of pract3
Included plan for pract3

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 \author{Bastiaan Weijers \\ 3256669 \and Colin Smits \\4075390}
@@
 First, we will discuss the techniques used in order to create a working system, and after that the requirements needed for the system to work will be explained.
+In addition to this paper, please refer to the film \footnote{\url{http://youtu.be/ye8yR4tcteo}}.
 
@@
 \section{Assumptions \& Requirements}
+\label{Ass}
 In order for our system to work we have to set the following requirements:
@@
 \section{Result}
+In the beginning, we struggled with implementing our idea, because it was difficult to detect skin color the way we wanted to (in HSV color space).
+However, when we started using YCbCr as color space, it worked wuite well, and we could continue.
+At last, we think that our product is overall robust (low amount of false positives) in a controlled environment (see Section \ref{Ass}).
 

### 2015-03-25 pract2/functions.py

@@ -52,7 +52,7 @@ def findDefects(frame, gray):
 			cnt = contours[n]
 			area = cv2.contourArea(cnt)
 					
-			if area > 10000:
+			if area > 4000:
 				ci.append(i)
 
 			n += 1
@@ -116,8 +116,8 @@ def getLengthY(vector):
 def getPreviousCenter(center, previous_centers):
 	i = 0
 	maxDistance = 10000000
-	print "Previous: %s" % (previous_centers)
-	print "Current: %s" % (str(center))
+	#print "Previous: %s" % (previous_centers)
+	#print "Current: %s" % (str(center))
 	for prev_center in previous_centers:
 
 		difference_X = center[0] - prev_center[0]
@@ -131,7 +131,7 @@ def getPreviousCenter(center, previous_centers):
 
 		i += 1
 
-	print "Related to center %s: %s = %s" % (str(center), current, previous_centers[current])
+	#print "Related to center %s: %s = %s" % (str(center), current, previous_centers[current])
 
 	return current	
 

### 2015-03-25 pract2/testBackground.py

@@ -5,20 +5,22 @@ import functions
 import win32com.client as comclt
 import win32api
 
-# Initialise Windows shell for communication
+# Initialiseer de Windows shell in Python zodat je naar de shell kunt scripten
 wsh= comclt.Dispatch("WScript.Shell")
 
 
 cap = cv2.VideoCapture(0)
-#fgbg = cv2.BackgroundSubtractorMOG()
+fgbg = cv2.BackgroundSubtractorMOG()
 
+
+#frame = cv2.imread("test2OtherLighting.jpg")
 #skin filter boundaries (tweak if necessary)
-lower = np.array([0, 140, 77], dtype = "uint8")
-upper = np.array([255, 173, 127], dtype = "uint8")
+lower = np.array([0, 45, 80], dtype = "uint8")
+upper = np.array([15, 255, 255], dtype = "uint8")
 handFound = False
 
 
-
+#"""
 #starting cam: no frame yet
 while (True):
     ret, frame1 = cap.read()
@@ -28,16 +30,15 @@ while (True):
 lastFrame = 0
 lastFrame2 = 0
 lastFound = 0
-findDelay = 30
+findDelay = 20
 maxDelay = 20
 oneHand = False
 twoHands = False
 trigger = 300
 _2HandsFound = False
 lastFound2 = 0
-framecount = 0
 
-trigger2Hands = 300
+trigger2Hands = 200
 
 #cam started, process frames
 while(cap.isOpened):
@@ -45,166 +46,135 @@ while(cap.isOpened):
 
     if ret == True:
 
-        # Delays in order to make sure a gesture will not be processed twice at once
-        
-
         if lastFound < findDelay:
             lastFound += 1
 
+        if lastFound2 < findDelay:
+            lastFound2 += 1
 
         if lastFrame2 < findDelay:
             lastFrame2 += 1
         else:
             twoHands = False
-            _2HandsFound = False
-
 
         img = frame.copy()
-
-        """
-        back = fgbg.apply(img)
-        img = cv2.bitwise_and(img, img, mask=back)
-        kernel = cv2.getStructuringElement(cv2.MORPH_ELLIPSE, (3, 3))
-        img = cv2.erode(img, kernel, iterations = 2)
-        img = cv2.dilate(img, kernel, iterations = 2)
-        """
-
-        # process the image to find skin locations
-        # Returns list: [ img, centers ] where centers = [(int, int)]
+       # back = fgbg.apply(img)
+       # img = cv2.bitwise_and(img, img, mask=back)
[…]

### 2015-03-25 pract2/testFace.py

@@ -1,7 +1,7 @@
 import numpy as np
 import cv2
 
-face_cascade = cv2.CascadeClassifier('haarcascade_frontalface_default.xml')
+face_cascade = cv2.CascadeClassifier('haarcascade_frontalface_alt2.xml')
 eye_cascade = cv2.CascadeClassifier('haarcascade_eye.xml')
 
 cap = cv2.VideoCapture(0)
@@ -21,8 +21,8 @@ while(cap.isOpened):
     if ret == True:
 		img = frame.copy()
 		gray = cv2.cvtColor(img, cv2.COLOR_BGR2GRAY)
-		faces = face_cascade.detectMultiScale(gray, 1.3, 5)
-		print str(faces)
+		faces = face_cascade.detectMultiScale(gray, 1.1, 5)
+		print len(faces)
 		for (x,y,w,h) in faces:
 		    cv2.rectangle(img,(x,y),(x+w,y+h),(255,0,0),2)
 		    roi_gray = gray[y:y+h, x:x+w]
@@ -31,7 +31,7 @@ while(cap.isOpened):
 		    for (ex,ey,ew,eh) in eyes:
 		        cv2.rectangle(roi_color,(ex,ey),(ex+ew,ey+eh),(0,255,0),2)
 
-		cv2.imshow('img',gray)
+		cv2.imshow('img',img)
      
     k = cv2.waitKey(30) & 0xff
     if k == 27:

### 2015-04-08 pract3/paper/paper.tex
first paper

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Bastiaan Weijers \\ 3256669 \and Colin Smits \\ 4075390}
\title{Evaluation of a Body Photo Browser}
\begin{document}
\maketitle
\abstract{Here comes the abstract}\\
\textbf{keywords:}

\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}
\section{Introduction}
\section{System Evaluation}
\section{User Evaluation}
\section{Discussion}

\thebibliography{99}

\end{document}

### 2015-04-08 pract3/plan/plan.tex
first paper

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Bastiaan Weijers \\ b.weijers@students.uu.nl \\ 3256669 \and Colin Smits \\ c.smits@uu.nl \\4075390}
\date{}
\title{\textbf{Plans for Assignment 3}}
\begin{document}
\maketitle
\section{Introduction}
The system we created is used to identify gestures in order to control a photo browser. The following gestures are defined as controls: swiping left and right will give the previous photo and next photo respectively. Right hand up, left hand down will rotate counter clockwise, and vice versa.

\section{System Evaluation}
In order to test the robustness, we will look at how many times the system correctly handles a gesture. For that, we can create a confusion matrix to compare the system's accuracy to what we defined to be true. We can specify the "ground truth" for this by using the definitions of the gestures given above. After that, we can measure the precision and recall. We think precision is most important, as we do not want false positives, during which our system will detect something when nothing occurred. We then want the average precision, so we will be using a PR-curve. In order to do this test correctly, we shall have to incorporate the notion ecological validity. Our setting thus will not be a laboratory, but a normal room with the same lighting during every test, and a background that is suitable for our system and does not change either.

\section{User Evaluation}
In order to get an idea of what the user thinks of our system, we can perform a certain amount of tests. As A/B testing and eye tracking cannot be performed since we only have one variation of the system and eye tracking is not important in this setting, we shall have to limit our evaluation to a usability interview. In this interview, we can start with putting the user in front of the system, and start asking what actions they would perform in order to use the system, and why. After that, we can introduce them the correct gestures, and ask what they think of that, and what they would change.

\end{document}

### 2015-04-08 pract3/paper/paper.tex
paper

@@
 \usepackage{graphicx}
+\usepackage{tikz}
+\usepackage{multirow}
+\usepackage{rotating}
+\usetikzlibrary{positioning}
 \author{Bastiaan Weijers \\ 3256669 \and Colin Smits \\ 4075390}
@@
 \section{System Evaluation}
+In this chapter we will discuss how the system evaluation was performed and what the results were of this evaluation. 
+\subsection{Methods and Environment}
+To quantify the performance of the Body Motion program, we setup a test environment to measure how well the system can translate user interaction to concrete system actions.
+The test setup is contained in an ecologically sound environment, in this case a bedroom with closed curtains was used with as a primary light source three led bulbs lighting the room.
+There was a single participant used to create quantifying data, whereas the particular user was well introduced to the workings, capabilities and limitations of the system.
+An analysis of how the system performs when used by lesser informed participants can be found later on in Section 3 - User Evaluation.
+\\ The software was used with a YCrCb spectrum at [..   ...] and a minimum delay of $70ms$ between actions, further settings were left at program-default.
+\\ For each action performed in the video recording, a chart was updated with the value the system interpreted it to be.
+Possible predicted and actual actions were: Next, Previous, CW rotation, CCW rotation and Nothing.
+Whereas "Nothing" is represented as an idle action wherein nothing happens.
+It is worth noting that this particular action can only be actively observed on the actual outputs of the system, as predicting idle actions is not possible in our system.
+\subsection{results}
+To achieve accurate and representing results, we have recorded a total of 159 actions that we will measure its interpretation by system off, this accounts to roughly 40 actions per possible move.
+The number was picked as such because a $n > 30$ will give us a normal distribution over our data [ref needed].
+All actions are recorded in a total of two video's, this allows the system to run in a continuous mode, without having to reinitialize after every action.
+This continuous mode should yield results that are more ecologically sound as opposed to having system global variables reinitialized with every action.
+\\ The evaluated output of a the actions will be tested against their expected outputs, the result of this can be found in Table~\ref{table:systemeval}.
+The data from the table shows the Predicted actions against the Actual actions delivered by the system for every possible action.
+From the table one can read how many predictions actually translated to correct behavior of the system, and how many did something unexpected.
+\\ Table~\ref{table:systemeval} can be parsed to a so-called Confusion Matrix, which is depicted in Table~\ref{table:confusionmatrix}.
+This matrix is a further generalization and will help us in the question on how well our system performs.
+Stated before (????, dit is wel in de presentatie gezegd maar nemen we 't op in de paper?), a preferred bias exists in having low false positives as this might be experienced by end-users as undesired behavior.
+From looking at the Confusion matrix we strongly suspect that we have satisfied this condition, but to confirm this we need to look at the Precision and Recall and establish a PR-curve.
+\\ hier een stuk over precision, recall en pr-curve.
+\\ moeten we nog een H0 opstellen voor de confusion matrix? We kunnen misschien proeven dat false positive < false negative?
+\\ Verder zien we uit Tabel 1 al dat de PR-curve voor swift-left/right er heel anders uit zal zien dan de PR-curve voor rotation.
+Die laatste is gewoon beroerd, terwijl de precision en recall voor left/right vrij goed zullen zijn.
+Dit apart testen denk ik in een subhypothese?
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\cline{3-7}
+\multicolumn{2}{c|}{} & \multicolumn{5}{c|}{Predicted action} \\
+\cline{3-7}
+\multicolumn{2}{c|}{} & Next & Previous & CW & CCW & Nothing \\
+\hline
+\multirow{6}{*}{\begin{sideways} Actual action ~ \end{sideways}} & Next & 34 & 0 & 0 & 0 & 0 \\
+\cline{2-7}
+& Previous & 1 & 32 & 0 & 0 & 2 \\
+\cline{2-7}
+& CW & 0 & 0  & 22  & 4  & 6  \\
+\cline{2-7}
+& CCW  & 0  & 0  & 0  & 14 & 0  \\
+\cline{2-7}
+& Nothing & 2 & 5  & 16  & 21  & 0  \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Testresults System Evaluation, $n=159$}
+\label{table:systemeval}
+\end{table}
+
+\par
+The data as presented can be further parsed into a confusion matrix which is shown in Table~\ref{table:confusionmatrix}.
+
+\begin{table}
+\begin{center}
+\begin{tikzpicture}[
+box/.style={draw,rectangle,minimum size=2cm,text width=1.5cm,align=left}]
+\matrix (conmat) [row sep=.1cm,column sep=.1cm] {
+\node (tpos) [box,
+    label=left:\( \mathbf{p'} \),
+    label=above:\( \mathbf{p} \),
+    ] {True \\ positive \\ 102};
+&
+\node (fneg) [box,
+    label=above:\textbf{n},
+    label=above right:\textbf{total},
+    label=right:\( \mathrm{P}' \)] {False \\ negative \\ 44};
+\\
+\node (fpos) [box,
+    label=left:\( \mathbf{n'} \),
+    label=below left:\textbf{total},
+    label=below:P] {False \\ positive \\ 5 };
+&
+\node (tneg) [box,
+    label=right:\( \mathrm{N}' \),
+    label=below:N] {True \\ negative \\ 0};
+\\
+};
+\node [left=.05cm of conmat,text width=1.5cm,align=right] {\textbf{actual \\ outcome}};
+\node [above=.05cm of conmat] {\textbf{prediction outcome}};
+
+\end{tikzpicture}
+\end{center}
+\label{table:confusionmatrix}
+\caption{Confusion Matrix}
+\end{table}
+
+\subsection{Discussion}
+Hier discussie over de resultaten.
+Hoewel de resultaten nog ontbreken, weten we waarschijnlijk al wel waar de resultation heengaan.
+Overall scores zullen 'mediocre' zijn at-best.
+Tenzij we de hypotheses opsplitsen en apart kijken naar swipe-left/right en rotatie.
+Dan kunnen we misschien iets zeggen over het verschil daarin.
+We kunnen dan ook "recommendations" maken.
+
 \section{User Evaluation}

### 2015-04-09 pract3/paper/paper.tex
paper verder, statistieken in excel

@@
 \author{Bastiaan Weijers \\ 3256669 \and Colin Smits \\ 4075390}
-\title{Evaluation of a Body Photo Browser}
+\title{Evaluation of a Body Photo Browser\footnote{Word Count:2957}}
+\renewcommand{\familydefault}{\sfdefault}
 \begin{document}
 \maketitle
-\abstract{Here comes the abstract}\\
-\textbf{keywords:}
+\abstract{A computer vision systems recognising gestures in order to operate a photo browser is reviewed.
+The system evaluation focuses mainly on the accuracy of the system.
+We see that some gestures are confused, limiting the system's overall performance.
+The user evaluation is carried out by usability questionnaires.
+An overall evaluation is given using the PSSUQ, which shows that the system does not meet the standards set for a computer vision system.
+Parts of the system are evaluated using ASQ and SEQ, which show difference between tasks within the system.
+Linking both, we see that the way the system is running short of recognising gestures influences the user's perception of the system.
+}
+
+\vspace{1cm}
+
+\noindent \textbf{Keywords:} Computer vision, gesture recognition, system evaluation, user evaluation
 
 \noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}
+
 \section{Introduction}
+The ease of using systems has become more important through the years.
+In order to check the usefulness, we need to evaluate the system.
+In this paper, we will discuss the usability of a body photo browser system, as will be defined later this section.
+The need for this is quite clear: the system is built for a user who wants a fluent experience, as well as a clear view of how to use the system.
+Taking this into consideration, we will discuss the system both from a system's perspective and the user's perspective.
+
+The purpose of this evaluation is mainly to identify pitfalls in the system's accuracy, as well as recognising limiting factors for a satisfying user experience.
+We want to know what adaptations must be made to the current system in order to create an overall better performing system in terms over ease, satisfaction and accuracy.
+
+In section 2, the system's reliability concerning robustness and accuracy will be discussed.
+After that, the usability of the system regarding ease of use and intuitiveness will be discussed in section 3.
+Last but not least, a conclusion will be drawn based upon the evaluations in sections 2 and 3.
+
+%INCLUDE research goals in introduction
+
+\subsection{The system}
+\label{sub:system}
+The system is a body photo browser. Using the system, one is able to scroll through and rotate pictures in the Windows Photo Gallery software %insert source.
+The following gestures are recognised by the system:
+
+\begin{description}
+\item[Swipe Left:] Using one hand (in view), going from right to left in the view. This will give the previous photo.
+\item[Swipe Right:] Using one hand (in view), going from left to right in the view. This will give the next photo.
+\item[Rotate left:] Using two hands (in view), going up with the right hand, and simultaneously going down with the left hand. This mimics the idea of rotating a picture counter clockwise.
+\item[Rotate Right:] Using two hands (in view), going up with the left hand, and simultaneously going down with the right hand. This quite mimics the idea of rotating a picture clockwise.
+\end{description}
+
+The system works by recognising skin colour, then tracking the hands blobs through the screen, and depending on the total offset measured, a gesture might or might not be recognised and handled.
+
 \section{System Evaluation}
 In this chapter we will discuss how the system evaluation was performed and what the results were of this evaluation.
+
 \subsection{Methods and Environment}
@@
 It is worth noting that this particular action can only be actively observed on the actual outputs of the system, as predicting idle actions is not possible in our system.
-\subsection{results}
+\subsection{Results}
 To achieve accurate and representing results, we have recorded a total of 159 actions that we will measure its interpretation by system off, this accounts to roughly 40 actions per possible move.
@@
 \section{User Evaluation}
+Now that we have had a look from a system point-of-view, we will now evaluate the system from a user perspective.
+First, the used methodology will be discussed.
+After that, the results and discussion will follow.
+
+\subsection{Methods and Environment}
+\label{sub:userMeth}
+In our user evaluation, we want to see whether our system is intuitive and easy to use.
+In order to do that, we need to have a number of people testing our system.
+First of all, we need to define the environment in which we will be testing.
+Using the notion of ecological validity, which states that the surroundings in which the research takes place must resemble a real-world setting, the tests will take place in normal rooms in terms of for example lighting and background.
+A laboratory would not be suitable as the user will react differently.
+The tests are then carried out as follows.
+
+%Brewer, M. (2000). Research Design and Issues of Validity. In Reis, H. and Judd, C. (eds) Handbook of Research Methods in Social and Personality Psychology. Cambridge:Cambridge University Press.)
+
+First of all, we ask the user to take a look at the system.
+We have defined four different gestures, and for each of the different options, we ask the user what would be his first opinion on how to perform the task.
+The user will give his own gesture which he would use in order to perform the action.
+That way, we can compare the thoughts of users to our own throughts on the used gestures.
+After that, we set up 10 scenarios, as shown in Table \ref{tab:scen}.
+The first four scenarios contain each of the different gestures in order to make the user comfortable with the system.
+After that, we start combining the gestures so that the user should be thinking about what gestures he must use in which order.
+Finally, we do not specify the gestures any more, but we describe the task by specifying pictures the user must recognise.
+In this way, we can test whether the user can perform the right actions while still thinking of what to do.
+The last part mainly shows how intuitive and easy our system can be used, as the user will be distracted from the gestures.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{r|l}
+\textbf{Scenario} & \textbf{Task} \\
+\hline
+1 & Navigate to the next photo \\ 
+2 & Navigate to the previous photo \\
+3 & Rotate the photo clockwise \\
+4 & Rotate the photo counter-clockwise \\
+5 & Navigate to the second next photo (twice to the next)\\ 
+6 & Rotate the previous photo clockwise \\
+7 & Rotate the photo and put it back up \\
+8 & Navigate to the photo containing [something] \\
+9 & Put the current (rotated) photo back up \\
+10 & Navigate to the rotated photo and put it back up \\
+\end{tabular}
+\end{center}
+\caption{\textit{Specified scenarios in the user tests}}
+\label{tab:scen}
+\end{table}
+
+Secondly, in order to get measurements, we set up a questionnaire.
+According to Sauro and Lewis \cite{Quantify}, and Tullis and Albert \cite{Measure}, there are different standardized questionnaires.
+For our system, we want to know whether the user is satisfied with the system as a whole.
+As defined by Sauro and Lewis, the PSSUQ (Post-study System Usability Questionnaire) is "designed to access users' perceived satisfaction with computer system applications" \cite{Quantify}.
+Looking at our goal, we think this questionnaire fits our research.
+Therefore, we incorporated this questionnaire at the end of the test, after all the scenarios have been completed.
+We chose the PSSUQ over the SUS (System Usablity Scale), which can also be used as a measure of the usability of a system, due to an increased reliability \cite{Quantify}.
+
+In addition, we thought that the user should be able to tell us what he perceived during each scenario.
+For that, we set up a combination of the ASQ (After Scenario Questionnaire) and the SEQ (Single Ease Question).
+The AQS was developed by Lewis in order to describe the user's satisfaction with completing a certain task \cite{Quantify}\cite{ASQ}.
+This can be perfectly used for our research goal.
+Moreover, the ASQ is linked to the PSSUQ, and also describes the user's satisfaction with a system, in this case a certain task.
+Furthermore, as the name suggests, the SEQ is used to describe a user's view on the ease of completing a task.
+The combination is used to get more reliable results.
+
+\subsection{Results}
+To start off, we look at the user's thoughts on the gestures to be used.
+In Table \ref{tab:compGes}, we show the percentage of user's defining a certain gesture for an action.
+We can see that for the navigation actions, the users differ in view, whereas for rotations, all the user say the same.
+\renewcommand{\arraystretch}{1.5}
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{l || l | c}
+\textbf{Action} & \textbf{ Users opinions} & \% \\
+\hline
+\multirow{2}{*}{Next photo} & Hand left to right & 75\\
+ & \textbf{Hand right to left }& \textbf{25} \\
+\hline
+\multirow{2}{*}{Previous photo} & \textbf{Hand left to right} & \textbf{25}\\
+& Hand right to left & 75\\
+\hline
+\multirow{2}{*}{Rotate left} & Rotate hands in circle counter-clockwise & 100 \\
+& \textbf{Left hand down, right hand up} & \textbf{0} \\
+\hline
+\multirow{2}{*}{Rotate Right} & Rotate hands in circle clockwise & 100 \\
+& \textbf{Left hand up, right hand down} & \textbf{0}
+\end{tabular}
+\end{center}
+\caption{\textit{Comparison of the implemented gestures (in \textbf{bold}) with the user's opinions on the gestures to be used. $ n = 4$}}
+\label{tab:compGes}
+\end{table}
+
+After that, we need to examine the results from the questionnaires. We will start by analysing the scenario questionnaires, the SEQ and ASQ.
+Firstly, the SEQ results are shown in Table \ref{tab:seq}.
+On first glance, we can see that for the scenarios containing only navigating to photos, the overall score is low ( $<$ 3), as opposed to those only containing rotations ( $>$4).
+Furthermore, the more complicated tasks show a broad confidence interval, meaning that there is much variation between users as far as ease of use is concerned.
+\\
+
+\textit{[ASQ results will follow in the final paper]\\}
+
+\renewcommand{\arraystretch}{1.0}
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{c  l || c || c | c}
+\textbf{Scenario} & \textbf{Description} & 
+\textbf{Lower} &
+\textbf{Mean} &
+\textbf{Upper} \\
+\hline
+1 & Navigate to the next photo & 0.58 & 1.5 & 2.42 \\
+2 & Navigate to the previous photo & 0.58 & 1.5 & 2.42 \\
+3 & Rotate the photo clockwise & 2.45 & 4.5 & 6.56 \\
+4 & Rotate the photo counter-clockwise  & 2.45 & 4.5 & 6.56 \\
+5 & Navigate to the second next photo & 0.73 & 2.25 & 3.77 \\
+6 & Rotate the previous photo clockwise & 0.40 & 3.0 & 5.60 \\
+7 & Rotate the photo and put it back up & 1.25 & 3.25 & 5.25\\
+8 & Navigate to the photo containing [something]& 1.25 & 3.25 & 5.25\\
+9 & Put the current (rotated) photo back up  & 0.10 & 3.0 & 5.90\\
+10 & Navigate to the rotated photo and put it back up & 0.70 & 2.0 & 3.30\\
+\end{tabular}
+\end{center}
+\caption{\textit{Results of the SEQ, showing the mean and 95\% confidence intervals for each of the scenarios. $n=4$}}
+\label{tab:seq}
+\end{table}
+
+Now we need to study the user's satisfaction overall. In Table \ref{tab:pssuq}, the results from the PSSUQ questionnaire are shown.
+
+\begin{table}[t!]
+\begin{center}
+\begin{tabular}{c || c || c | c}
+\textbf{Question} &
+\textbf{Lower} &
+\textbf{Mean} &
+\textbf{Upper} \\
+\hline
+1  & 1.95 & 2.75 & 3.55 \\
+2 & 1.23 & 2.75 & 4.27 \\
+3 & 1.95 & 2.75 & 3.55\\
+4  & 0.75 & 3 & 5.25 \\
+5 & 1.23 & 2.75 & 4.27 \\
+6 & 1.73 & 3.25 & 4.77 \\
+7 & 3.36 & 5.75 & 8.14\\
+8 & 2.70 & 4 & 5.30\\
+9  & 1.74 & 4.5 & 7.26\\
+10 & 2.75 & 5 & 7.25\\
+11 & 2.36 & 4.75 & 7.14 \\
+12 & 2.75 & 5 & 7.25\\
+13 & 0.75 & 3 & 5.25\\
+14 & 2.46 & 3.25 & 4.05\\
+15 & 2.58 & 3.5 & 4.42\\
+16 & 1.73 & 3.25 & 4.77 \\
+\hline 
+\hline
+Avg. 1-6 & 1.47 & 2.88 & 4.28 \\
+Avg. 7-12 & 2.61 & 4.83 & 7.05 \\
+Avg. 13-15  & 1.92 & 3.25  & 4.57 \\
+Avg. 1-16  & 2.00 & 3.70 &  5.40 \\
+\end{tabular} 
+\caption{PSSUQ results showing the mean for each question compared to the 95\% confidence intervals. $n=4$}
+\label{tab:pssuq}
+\end{center}
+\end{table}
+
+\subsection{Discussion}
+From the results obtained for the user's opinion on the gestures, we can conclude that the navigation is something that matters from person to person.
+The intention of the implemented gesture was to simulate a swipe on, for example, a tablet.
+To go to the next photo, one must then swipe from right to left, seemingly pushing the photo roll to the left.
+However, users perceive that it could also be seen as the direction to where one wants to go.
+In this case, the meaning of a swipe to the left is that the photo to the left should be shown, thus the previous.
+This is something that has to be worked out further.
+Another striking observation is that users all have the same opinion on rotation.
+However, this differs from our own implementation.
+Looking closely to the implementation, however, we see that the gesture implemented resembles a rotating movement without circulation.
+The circular limitation still hinders smooth use of the system, as can also be seen in the SEQ.
+
+The SEQ shows us that the system's easy lies in the navigation, and that the rotation is more difficult.
+As stated before, the cause of this might be the implementation of the gestures as is.
+The navigation issue is resolved due to the fact that the right gestures are explained after the first questions (Section \ref{sub:userMeth}.
+
+From the PSSUQ, we can conclude that the system's performance is not good enough, when compared to the PSSUQ norms \cite{Quantify}.
+
+Overall, for the user evaluation, we can generally say that when only navigation is incorporated, the user is able to do tasks with ease.
+On the contrary, when rotation is included, the user will start to struggle.
+Nonetheless, we still have to be careful concluding from our result set, as the sample size is much too low ($n=4$) in order to get a clear representation.
+
 \section{Discussion}
+All in all, when we link the system evaluation to the user evaluation, we can see one striking property already.
+Due to problems with the implementation of the rotating gestures, the system performs poor and the user cannot use the system according to plan.
+
+In the system evaluation, we have seen that rotating gestures are easily confused with the other rotation, and that the system sometimes does not recognise any gesture at all when rotating.
+In the user evaluation, we have seen these properties return in the SEQ, where scenarios involving rotation were seen less satisfactory to the user.
+This implies that for further research, the system's implementation of rotation must be reviewed and edited, in order to get a more accurate and more satisfying system.
 
 \thebibliography{99}
+\bibitem{Quantify} J. Sauro, J. R. Lewis: \textit{"Quantifying the User Experience"}, Elsevier, 2012
+\bibitem{Measure} T. Tullis, W. Albert: \textit{"Measuring the User Experience: Collecting, Analyzing, and Presenting Usability Metrics" (2nd Edition)}, Elsevier, 2013 
+\bibitem{ASQ} J.
+R.
+Lewis: \textit{"IBM computer usablity satisfaction questionnaires: Psychometric evaluation and instructions for use"}, International Journal for Human Computer Interactions, 7 (1995), 57-78
+\bibitem{EcoVal} M.
+Brewer: \textit{"Research Design and Issues of Validity"}.
+In: H.
+Reis, C.
+Judd (eds), \textit{"Handbook of Research Methods in Social and Personality Psychology"}, Cambridge University Press, 2000
 

### 2015-04-16 pract3/paper/paper.tex

@@
 \usepackage{rotating}
+\usepackage{cite}
 \usetikzlibrary{positioning}
+
 \author{Bastiaan Weijers \\ 3256669 \and Colin Smits \\ 4075390}
@@
 An analysis of how the system performs when used by lesser informed participants can be found later on in Section 3 - User Evaluation.
-\\ The software was used with a YCrCb spectrum at [..   ...] and a minimum delay of $70ms$ between actions, further settings were left at program-default.
+\\ The software was used with a YCrCb spectrum at [180 - 210] and a minimum delay of $70ms$ between actions, further settings were left at program-default.
 \\ For each action performed in the video recording, a chart was updated with the value the system interpreted it to be.
@@
 First of all, we need to define the environment in which we will be testing.
-Using the notion of ecological validity, which states that the surroundings in which the research takes place must resemble a real-world setting, the tests will take place in normal rooms in terms of for example lighting and background.
+Using the notion of ecological validity, which states that the surroundings in which the research takes place must resemble a real-world setting\cite{brewer2000research}, the tests will take place in normal rooms in terms of for example lighting and background.
 A laboratory would not be suitable as the user will react differently.
 The tests are then carried out as follows.
-
-%Brewer, M. (2000). Research Design and Issues of Validity. In Reis, H. and Judd, C. (eds) Handbook of Research Methods in Social and Personality Psychology. Cambridge:Cambridge University Press.)
 
@@
 Secondly, in order to get measurements, we set up a questionnaire.
-According to Sauro and Lewis \cite{Quantify}, and Tullis and Albert \cite{Measure}, there are different standardized questionnaires.
+According to Sauro and Lewis \cite{sauro2012quantifying}, and Tullis and Albert \cite{albert2013measuring}, there are different standardized questionnaires.
 For our system, we want to know whether the user is satisfied with the system as a whole.
-As defined by Sauro and Lewis, the PSSUQ (Post-study System Usability Questionnaire) is "designed to access users' perceived satisfaction with computer system applications" \cite{Quantify}.
+As defined by Sauro and Lewis, the PSSUQ (Post-study System Usability Questionnaire) is "designed to access users' perceived satisfaction with computer system applications" \cite{sauro2012quantifying}.
 Looking at our goal, we think this questionnaire fits our research.
 Therefore, we incorporated this questionnaire at the end of the test, after all the scenarios have been completed.
-We chose the PSSUQ over the SUS (System Usablity Scale), which can also be used as a measure of the usability of a system, due to an increased reliability \cite{Quantify}.
+We chose the PSSUQ over the SUS (System Usablity Scale), which can also be used as a measure of the usability of a system, due to an increased reliability \cite{sauro2012quantifying}.
 
@@
 For that, we set up a combination of the ASQ (After Scenario Questionnaire) and the SEQ (Single Ease Question).
-The AQS was developed by Lewis in order to describe the user's satisfaction with completing a certain task \cite{Quantify}\cite{ASQ}.
+The AQS was developed by Lewis in order to describe the user's satisfaction with completing a certain task \cite{sauro2012quantifying}\cite{lewis1995ibm}.
 This can be perfectly used for our research goal.
@@
 
-From the PSSUQ, we can conclude that the system's performance is not good enough, when compared to the PSSUQ norms \cite{Quantify}.
+From the PSSUQ, we can conclude that the system's performance is not good enough, when compared to the PSSUQ norms \cite{sauro2012quantifying}.
 
@@
 
-\thebibliography{99}
-\bibitem{Quantify} J. Sauro, J. R. Lewis: \textit{"Quantifying the User Experience"}, Elsevier, 2012
-\bibitem{Measure} T. Tullis, W. Albert: \textit{"Measuring the User Experience: Collecting, Analyzing, and Presenting Usability Metrics" (2nd Edition)}, Elsevier, 2013 
-\bibitem{ASQ} J.
-R.
-Lewis: \textit{"IBM computer usablity satisfaction questionnaires: Psychometric evaluation and instructions for use"}, International Journal for Human Computer Interactions, 7 (1995), 57-78
-\bibitem{EcoVal} M.
-Brewer: \textit{"Research Design and Issues of Validity"}.
-In: H.
-Reis, C.
-Judd (eds), \textit{"Handbook of Research Methods in Social and Personality Psychology"}, Cambridge University Press, 2000
+\bibliography{mybib}{}
+\bibliographystyle{plain}
 

### 2015-04-25 pract3/paper/paper.tex
system eval af? beetje matig verhaal, waardig voor een zesje

@@
 There was a single participant used to create quantifying data, whereas the particular user was well introduced to the workings, capabilities and limitations of the system.
-An analysis of how the system performs when used by lesser informed participants can be found later on in Section 3 - User Evaluation.
-\\ The software was used with a YCrCb spectrum at [180 - 210] and a minimum delay of $70ms$ between actions, further settings were left at program-default.
+5\\ The software was used with a YCrCb spectrum at [180 - 210] and a minimum delay of $70ms$ between actions, further settings were left at program-default.
 \\ For each action performed in the video recording, a chart was updated with the value the system interpreted it to be.
@@
 It is worth noting that this particular action can only be actively observed on the actual outputs of the system, as predicting idle actions is not possible in our system.
+%\par
+%Furthermore, we go ahead and compare our data with data we gained as byproduct from the User Evaluation (Section 3).
+The comparison of these datasets will give an indication how representative the controlled environment for the system evaluation was.
+
 \subsection{Results}
 To achieve accurate and representing results, we have recorded a total of 159 actions that we will measure its interpretation by system off, this accounts to roughly 40 actions per possible move.
-The number was picked as such because a $n > 30$ will give us a normal distribution over our data [ref needed].
 All actions are recorded in a total of two video's, this allows the system to run in a continuous mode, without having to reinitialize after every action.
 This continuous mode should yield results that are more ecologically sound as opposed to having system global variables reinitialized with every action.
-\\ The evaluated output of a the actions will be tested against their expected outputs, the result of this can be found in Table~\ref{table:systemeval}.
-The data from the table shows the Predicted actions against the Actual actions delivered by the system for every possible action.
-From the table one can read how many predictions actually translated to correct behavior of the system, and how many did something unexpected.
-\\ Table~\ref{table:systemeval} can be parsed to a so-called Confusion Matrix, which is depicted in Table~\ref{table:confusionmatrix}.
-This matrix is a further generalization and will help us in the question on how well our system performs.
-Stated before (????, dit is wel in de presentatie gezegd maar nemen we 't op in de paper?), a preferred bias exists in having low false positives as this might be experienced by end-users as undesired behavior.
-From looking at the Confusion matrix we strongly suspect that we have satisfied this condition, but to confirm this we need to look at the Precision and Recall and establish a PR-curve.
-\\ hier een stuk over precision, recall en pr-curve.
-\\ moeten we nog een H0 opstellen voor de confusion matrix? We kunnen misschien proeven dat false positive < false negative?
-\\ Verder zien we uit Tabel 1 al dat de PR-curve voor swift-left/right er heel anders uit zal zien dan de PR-curve voor rotation.
-Die laatste is gewoon beroerd, terwijl de precision en recall voor left/right vrij goed zullen zijn.
-Dit apart testen denk ik in een subhypothese?
-
 \begin{table}
@@
 \end{table}
+\\ The evaluated output of a the actions will be tested against their expected outputs, the result of this can be found in Table~\ref{table:systemeval}.
+The data from the table shows the Predicted actions against the Actual actions delivered by the system for every possible action.
+From the table one can read how many predictions actually translated to correct behavior of the system, and how many did something unexpected.
+\\ Table~\ref{table:systemeval} can be parsed to a so-called Confusion Matrix, which is depicted in Table~\ref{table:confusionmatrix}.
+This matrix is a further generalization and will help us in the question on how well our system performs.
+With a gesture controlled interface like the one tested, a preferred bias exists in having low false positives, possibly even in spite of an increased number of false negatives.
+From looking at the Confusion matrix we strongly suspect that we have satisfied this condition, but to confirm this we need to look at the Precision, Recall and Accuracy, shown in Table~\ref{table:systemevalvars}.
+\par The system seems to perform mediocre if it comes down to the Accuracy and Recall, but it is quite $Precise$.
+
+\begin{table}[!h]
+\centering
+\begin{tabular}{lll}
+\textbf{Accuracy}          & \textbf{Recall}            & \textbf{Precision}         \\
+\multicolumn{1}{c}{67,5\%} & \multicolumn{1}{c}{69,9\%} & \multicolumn{1}{c}{95,3\%}
+\end{tabular}
+\caption{Accuracy, Prevision and Recall of System Evaluation}
+\label{table:systemevalvars}
+\end{table}
 
@@
 
-\begin{table}
+\begin{table}[h]
 \begin{center}
@@
 \subsection{Discussion}
-Hier discussie over de resultaten.
-Hoewel de resultaten nog ontbreken, weten we waarschijnlijk al wel waar de resultation heengaan.
-Overall scores zullen 'mediocre' zijn at-best.
-Tenzij we de hypotheses opsplitsen en apart kijken naar swipe-left/right en rotatie.
-Dan kunnen we misschien iets zeggen over het verschil daarin.
-We kunnen dan ook "recommendations" maken.
+Having looked at and quantified obtained data, we now have a good idea of how well the system performs.
+One of the initial preference biases was the preference of a high Precision.
+Clearly, from look at Table~\ref{table:systemevalvars}, we see that the Precision is very high.
+However, the high precision comes at a price of having a mediocre Recall.
+Perhaps one might argue that the system was designed on forehand to prefer higher precision at the cost of yielding a lower recall, but the results don't lie and give a total of 67,5\% on the overall Accuracy.
+\par From Table~\ref{table:confusionmatrix}, the confusion matrix, it seems that there exists a certain center of gravity if it comes down to where errors are produced.
+Namely CW/CCW rotations produce almost as much errors as successes.
+In this system evaluation an equal distribution over the actions was chosen, but perhaps there exists some difference in the usage of actions in the real world.
+So it could be that while Rotations generate a lot of errors, the actual usage of rotations by end-users might be low and thus resulting in a higher recall and accuracy.
 

### 2015-04-28 pract3/paper/paper.tex
paper af

@@
 
-\author{Bastiaan Weijers \\ 3256669 \and Colin Smits \\ 4075390}
-\title{Evaluation of a Body Photo Browser\footnote{Word Count:2957}}
+\author{Bastiaan Weijers \\ B.Weijers@students.uu.nl \\ 3256669 \and Colin Smits \\ C.Smits@uu.nl \\ 4075390}
+\title{Evaluation of a Body Photo Browser\footnote{Word Count:3698}}
 \renewcommand{\familydefault}{\sfdefault}
@@
 \maketitle
-\abstract{A computer vision systems recognising gestures in order to operate a photo browser is reviewed.
+\abstract{\noindent A computer vision systems recognising gestures in order to operate a photo browser is reviewed.
 The system evaluation focuses mainly on the accuracy of the system.
@@
 The purpose of this evaluation is mainly to identify pitfalls in the system's accuracy, as well as recognising limiting factors for a satisfying user experience.
-We want to know what adaptations must be made to the current system in order to create an overall better performing system in terms over ease, satisfaction and accuracy.
+We want to know what adaptations must be made to the current system in order to create an overall better performing system in terms of ease, satisfaction and accuracy.
 
 In section 2, the system's reliability concerning robustness and accuracy will be discussed.
-After that, the usability of the system regarding ease of use and intuitiveness will be discussed in section 3.
+For this, we will look at the accuracy of the system using a confusion matrix in order to get the system's precision and recall.
+After that, the usability of the system regarding ease of use and intuitiveness will be discussed in section 3 by the use of standard questionnaires (PSSUQ, ASQ and SEQ, see \ref{sub:userMeth}).
 Last but not least, a conclusion will be drawn based upon the evaluations in sections 2 and 3.
@@
 \end{tabular}
-\caption{Accuracy, Prevision and Recall of System Evaluation}
+\caption{Accuracy, Precision and Recall of System Evaluation}
 \label{table:systemevalvars}
@@
 For that, we set up a combination of the ASQ (After Scenario Questionnaire) and the SEQ (Single Ease Question).
-The AQS was developed by Lewis in order to describe the user's satisfaction with completing a certain task \cite{sauro2012quantifying}\cite{lewis1995ibm}.
+The ASQ was developed by Lewis in order to describe the user's satisfaction with completing a certain task \cite{sauro2012quantifying}\cite{lewis1995ibm}.
 This can be perfectly used for our research goal.
@@
 In Table \ref{tab:compGes}, we show the percentage of user's defining a certain gesture for an action.
-We can see that for the navigation actions, the users differ in view, whereas for rotations, all the user say the same.
+We can see that for the navigation actions, the users differ in view, whereas for rotations, all the users say the same.
 \renewcommand{\arraystretch}{1.5}
@@
 \hline
-\multirow{2}{*}{Next photo} & Hand left to right & 75\\
- & \textbf{Hand right to left }& \textbf{25} \\
+\multirow{2}{*}{Next photo} & Hand left to right & 43 \\
+ & \textbf{Hand right to left }& \textbf{57} \\
 \hline
-\multirow{2}{*}{Previous photo} & \textbf{Hand left to right} & \textbf{25}\\
-& Hand right to left & 75\\
+\multirow{2}{*}{Previous photo} & \textbf{Hand left to right} & \textbf{57}\\
+& Hand right to left & 43\\
 \hline
@@
 \end{center}
-\caption{\textit{Comparison of the implemented gestures (in \textbf{bold}) with the user's opinions on the gestures to be used. $ n = 4$}}
+\caption{\textit{Comparison of the implemented gestures (in \textbf{bold}) with the user's opinions on the gestures to be used. $ n = 7$}}
 \label{tab:compGes}
@@
 \\
-
-\textit{[ASQ results will follow in the final paper]\\}
 
@@
 \hline
-1 & Navigate to the next photo & 0.58 & 1.5 & 2.42 \\
-2 & Navigate to the previous photo & 0.58 & 1.5 & 2.42 \\
-3 & Rotate the photo clockwise & 2.45 & 4.5 & 6.56 \\
-4 & Rotate the photo counter-clockwise  & 2.45 & 4.5 & 6.56 \\
-5 & Navigate to the second next photo & 0.73 & 2.25 & 3.77 \\
-6 & Rotate the previous photo clockwise & 0.40 & 3.0 & 5.60 \\
-7 & Rotate the photo and put it back up & 1.25 & 3.25 & 5.25\\
-8 & Navigate to the photo containing [something]& 1.25 & 3.25 & 5.25\\
-9 & Put the current (rotated) photo back up  & 0.10 & 3.0 & 5.90\\
-10 & Navigate to the rotated photo and put it back up & 0.70 & 2.0 & 3.30\\
+1 & Navigate to the next photo & 0.93 & 2.43 & 3.93 \\
+2 & Navigate to the previous photo & 0.93 & 2.43 & 3.93 \\
+3 & Rotate the photo clockwise & 1.84 & 3.43 & 5.02 \\
+4 & Rotate the photo counter-clockwise  & 1.84 & 3.43 & 5.02 \\
+5 & Navigate to the second next photo & 1.59 & 2.29 & 2.98 \\
+6 & Rotate the previous photo clockwise & 1.55 & 2.71 & 3.87 \\
+7 & Rotate the photo and put it back up & 1.55 & 2.71 & 3.87\\
+8 & Navigate to the photo containing [something]& 1.39 & 2.57 & 3.75\\
+9 & Put the current (rotated) photo back up  & 1.03 & 2.42 & 3.83\\
+10 & Navigate to the rotated photo and put it back up & 1.41 & 2.28 & 3.17\\
 \end{tabular}
 \end{center}
-\caption{\textit{Results of the SEQ, showing the mean and 95\% confidence intervals for each of the scenarios. $n=4$}}
+\caption{\textit{Results of the SEQ, showing the mean and 95\% confidence intervals for each of the scenarios. $n=7$}}
 \label{tab:seq}
@@
 
-Now we need to study the user's satisfaction overall. In Table \ref{tab:pssuq}, the results from the PSSUQ questionnaire are shown.
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{l | llllllllll}
+            & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8} & \textbf{9} & \textbf{10} \\
+\hline
+\textbf{1}  & 1.00       & -          & -          & -          & -          & -          & -          & -          & -          & -           \\
+\textbf{2}  & 0.69        & 1.00       & -          & -          & -          & -          & -          & -          & -          & -           \\
+\textbf{3}  & 0.76       & 1.00       & 1.00       & -          & -          & -          & -          & -          & -          & -           \\
+\textbf{4}  & -0.97      & -0.5       & -0.58      & 1.00       & -          & -          & -          & -          & -          & -           \\
+\textbf{5}  & 0.76       & 1.00       & 1.00       & -0.58      & 1.00       & -          & -          & -          & -          & -           \\
+\textbf{6}  & 0.69       & 1.00       & 1.00       & -0.50      & 1.00       & 1.00       & -          & -          & -          & -           \\
+\textbf{7}  & 0.33       & 0.91       & 0.87       & -0.09      & 0.87       & 0.91       & 1.00       & -          & -          & -           \\
+\textbf{8}  & 1.00       & 0.69       & 0.76       & -0.97      & 0.76       & 0.69       & 0.33       & 1.00       & -          & -           \\
+\textbf{9}  & 0.50       & -0.28      & -0.19      & -0.69      & -0.19      & -0.28      & -0.28      & 0.50       & 1.00       &             \\
+\textbf{10} & 0.50       & 0.97       & 0.95       & -0.28      & 0.94       & 0.97       & 0.97       & 0.50       & -0.50      & 1.00       
+\end{tabular}
+\caption{\textit{Results of the ASQ, showing the correlation of overall usability between scenarios, $n=7$}}
+\label{tab:ASQ}
+\end{center}
+\end{table}
+
+The results from the ASQ are depicted in Table \ref{tab:ASQ}.
+Each cell represents the correlation of the scenarios as obtained after doing a factor analysis on the data gotten.
+Striking is the fact that scenarios 2, 3, 5 and 6 are tightly correlated in terms of ease.
+However, in comparison with scenario 1, which is opposite of scenario 2, these scenarios do not have a strong correlation.\\
+
+Now we need to study the user's satisfaction overall.
+In Table \ref{tab:pssuq}, the results from the PSSUQ questionnaire are shown.
+For each question, the mean of the scores is shown as well as the values for the 95\% confidence intervals.
+At the bottom, the averages per group of questions is shown, in compliance with the grouping of questions according to Sauro and Lewis \cite{sauro2012quantifying}.
 
@@
 \hline
-1  & 1.95 & 2.75 & 3.55 \\
-2 & 1.23 & 2.75 & 4.27 \\
-3 & 1.95 & 2.75 & 3.55\\
-4  & 0.75 & 3 & 5.25 \\
-5 & 1.23 & 2.75 & 4.27 \\
-6 & 1.73 & 3.25 & 4.77 \\
-7 & 3.36 & 5.75 & 8.14\\
-8 & 2.70 & 4 & 5.30\\
-9  & 1.74 & 4.5 & 7.26\\
-10 & 2.75 & 5 & 7.25\\
-11 & 2.36 & 4.75 & 7.14 \\
-12 & 2.75 & 5 & 7.25\\
-13 & 0.75 & 3 & 5.25\\
-14 & 2.46 & 3.25 & 4.05\\
-15 & 2.58 & 3.5 & 4.42\\
-16 & 1.73 & 3.25 & 4.77 \\
+1 & 2.08 & 2.57 & 3.06 \\
+2 & 1.70 & 2.43 & 3.15 \\
+3 & 2.26 & 2.71 & 3.17\\
+4 & 1.87 & 2.86 & 3.85 \\
+5 & 0.93 & 2.00 & 3.07 \\
+6 & 2.84 & 3.57 & 4.30 \\
+7 & 4.73 & 5.86 & 6.98\\
+8 & 1.90 & 3.14 & 4.39\\
+9  & 2.31 & 4.29 & 6.26\\
+10 & 2.31 & 4.00 & 5.69\\
+11 & 1.59 & 3.43 & 5.27\\
+12 & 3.40 & 4.86 & 6.31\\
+13 & 1.44 & 2.71 & 3.99\\
+14 & 2.22 & 2.86 & 3.99\\
+15 & 2.25 & 3.43 & 4.61\\
+16 & 2.03 & 2.86 & 3.69 \\
 \hline 
 \hline
-Avg. 1-6 & 1.47 & 2.88 & 4.28 \\
-Avg. 7-12 & 2.61 & 4.83 & 7.05 \\
-Avg. 13-15  & 1.92 & 3.25  & 4.57 \\
-Avg. 1-16  & 2.00 & 3.70 &  5.40 \\
+Avg. 1-6 & 1.95 & 2.69 & 3.43 \\
+Avg. 7-12 & 2.71 & 4.26 & 5.82 \\
+Avg. 13-15  & 1.97 & 3.00  & 4.03 \\
+Avg. 1-16  & 2.24 & 3.35 &  4.46 \\
 \end{tabular} 
-\caption{PSSUQ results showing the mean for each question compared to the 95\% confidence intervals. $n=4$}
+\caption{PSSUQ results showing the mean for each question compared to the 95\% confidence intervals. $n=7$}
 \label{tab:pssuq}
@@
 \end{table}
+
+\pagebreak
 
@@
 This is something that has to be worked out further.
+
 Another striking observation is that users all have the same opinion on rotation.
@@
 
-The SEQ shows us that the system's easy lies in the navigation, and that the rotation is more difficult.
+The SEQ results hint that the system's ease lies in the navigation, and that the rotation is more difficult.
+We can test this using a t-test on paired samples on the obtained results for the SEQ.
+Scenarios 3 and 4 are more difficult than scenarios 1 and 2 (M = 2, SD = 1.15).
+This difference is statistically significant: t(6) = 4.58, p $<$ 0.05, two-tailed.
+
  As stated before, the cause of this might be the implementation of the gestures as is.
@@
 
+In addition to the results from the SEQ, the ASQ shows us a difference between the rotations and the navigations in terms of correlation.
+However, we need to mention that there are inconsistencies in the data too, for example when comparing scenarios 1 and 2: the only changed factor is the direction in which navigation must be performed.
+However, we see that there is a low correlation between these two scenarios, whereas scenario 2 and 3, which are navigation as opposed to rotation, have a high correlation.
+
 From the PSSUQ, we can conclude that the system's performance is not good enough, when compared to the PSSUQ norms \cite{sauro2012quantifying}.
+The obtained values for each of the questions are slightly higher than wanted.
 
@@
 On the contrary, when rotation is included, the user will start to struggle.
-Nonetheless, we still have to be careful concluding from our result set, as the sample size is much too low ($n=4$) in order to get a clear representation.
+Nonetheless, we still have to be careful concluding from our result set, as the sample size is much too low ($n=7$) in order to get a clear representation.
 
 \section{Discussion}
-All in all, when we link the system evaluation to the user evaluation, we can see one striking property already.
-Due to problems with the implementation of the rotating gestures, the system performs poor and the user cannot use the system according to plan.
+To round off, when looking at the system evaluation, we see that the systems performance in terms of accuracy is quite well.
+Having a high precision comes with the cost of a decreased recall.
+Still, the accuracy amounts to a steady 67.5\%.
+Continuing on this, we have seen that the systems confuses the rotation gestures greatly.
+Sometimes the rotations are not even recognised by the system as such, causing erratic behaviour.
 
-In the system evaluation, we have seen that rotating gestures are easily confused with the other rotation, and that the system sometimes does not recognise any gesture at all when rotating.
-In the user evaluation, we have seen these properties return in the SEQ, where scenarios involving rotation were seen less satisfactory to the user.
-This implies that for further research, the system's implementation of rotation must be reviewed and edited, in order to get a more accurate and more satisfying system.
+The user evaluation shows the same problems. Where navigation is seemingly easy to do, rotations are experienced as much more difficult to do overall.
+
+Linking these two, we can conclude that the implementation of the rotations is hindering a smooth use of the system.
+The problems arising from the system not recognising a significant amount of gestures causes users to perform less on some of the tasks set.
+On the other hand, the way navigation is implemented is simple: the system can handle the gestures easily, and the user is satisfied by the way he can browse through the photos.
+This, however, needs to be said with caution, as users have a different opinion on what direction (left-right or right-left) is linked to a certain navigational task (next / previous).
+
+Taking this into account, we can safely say that further examination of the implementation of the rotations has to be done carefully.
+By improving the system's performance on this type of gestures, we can maybe improve the user's perception of the system, as we can conclude that these are linked.
+
+In this evaluation, we have not looked at the boundaries of the system in terms of lighting, difference between user's, and so on.
+Therefore, more research could be performed on the influence of colors and lighting on our system's performance.
+After examination of the gestures, we can perform A/B-testing in order to see whether the newly set gesture is good enough.
 
@@
 
+\section*{Acknowledgement:}
+\textit{We would like to thank each and every one of the willing volunteers for participating in our evaluation.} 
 \end{document}
